---
title: "R Notebook"
output: html_notebook
---
```{r}
library(readxl)
#library()
dta<-read_excel("data/mbta.xlsx", skip = 1,range=cell_cols(2:60))
View(dta)
```

This is used to unpivot row header to columns
```{r}
library(dplyr)
library(tidyr)
dta_tidy <-dta%>% gather(`2007-01`:`2011-10`, 
                          key="year", value="passengers")
View(dta_tidy)
```

 separate into separate columns key:value pairs
```{r}
dta_tidy <- dta_tidy %>% separate(year, into = c("year", "month"))
View(dta_tidy)
```


```{r}
dta_tidy<-dta_tidy%>%spread(mode, passengers)
View(dta_tidy)
```

subset rows
```{r}
dta_tidy <- dta_tidy %>% .[,c(1:2,6:8)]
```

```{r}
dta_tidy
```
install hflights datasets from R repository
```{r}
library(hflights)
data("hflights")
View(hflights)
```

Using the filter function
```{r}
filter(hflights, Year==2011, Month==2)
```
```{r}
fil <-filter(hflights, Dest%in%c("FLL", "IAH"))
fil
```
```{r}
dist <- filter(hflights, between(Distance, 224,944))[,11:16]
dist
```
### Tasks


1. All flights in April 2011

```{r}
april_dept<-filter(hflights,Year==2011, Month==4)
april_dept
```
2. All flights in AA & WN
```{r}
aw_flights = filter(hflights, UniqueCarrier%in%c("AA", "WN"))
aw_flights
```
SOrt values
By default, arrange sort vlaues ascending order. This is why we have to use desc

```{r}
sort_flight = arrange(hflights, desc(hflights$ArrDelay))
sort_flight
```

OR
```{r}
sort_flight = arrange(hflights, desc(ArrDelay))
sort_flight
```
using select function
the comma is needed so as to return only 4 rows
```{r}
select_rows <- select(hflights, Year, Month, FlightNum, AirTime)[1:4,]
select_rows
```
using Select helper functions
```{r}
vars <-c("Year", "Month", "DayofMonth", "DayofMonth", "ArrTIme")
s_1 <- select(hflights, one_of(vars))
head(s_1)[1:6,]
```
Using the mutate function
```{r}
hflights%>%
  select(ends_with("Delay"), Distance, AirTime)%>%
  mutate(time_gain = ArrDelay - DepDelay,
         speed = Distance/AirTime*60)
```


```{r}
hflights %>%
  ggroup_by(Yeear, Month, DayofMonth)%>%
  summarise(delay=mean(DepDelay, na.rm = TRUE))
```

