---
title: "Week2"
author: "Patrick"
date: "7/10/2020"
output: html_document
---

## Loading Data into Files
Loading data into R without using file.chose()
In this case, the data must be in the same path as this .Rmd file
```{r}
who_suicide <- read.table('~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/who_suicide_statistics.csv',sep = ",", header =T)
head(who_suicide)
```
```{r}
auto <- read.table('~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/auto.txt',sep = ",", header =T)
head(auto)
```
```{r}
potato <- read.table('~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/potatoes.txt', sep="\t", header=T)
head(potato)
```
#### read.csv2() for reading csv files
`Note:` read.csv2() is used to read only comma separated data and does not use the parameter sep

```{r}
other_csv <- read.csv2("~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/who_suicide_statistics_semi.csv", header = T)

head(other_csv)
```
#### using read.delim() for tsv files
`NB:` read.delim() works for tab separated files
```{r}
delim <- read.delim('~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/potatoes.txt', header=F)
head(delim)
```
#### package readr
a package called readr() is 10x faster but it does not come default with R. It has been installed unlike the base functions
Base functions for reading file include:
- read.delim()
- read.table()
-read.csV2()

#### Importing excel files 
> read_excel() is not a base R function. It has to be imported into r as a package.
> `NB:` The col_names here also refers to header as in other file reader functions
> sheet paramter refers to the shhet we want to load. We cannot load more than one sheet at a time

```{r}
library(readxl)
```
```{r}
sht1 <-read_excel('~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/Employees.xlsx', col_names = T, sheet = 1)
head(sht1)
```


