---
title: "Week2"
author: "Patrick"
date: "7/10/2020"
output: html_document
---

## Loading Data into Files
Loading data into R without using file.chose()
In this case, the data must be in the same path as this .Rmd file
```{r}
who_suicide <- read.table('~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/who_suicide_statistics.csv',sep = ",", header =T)
head(who_suicide)
```
```{r}
auto <- read.table('~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/auto.txt',sep = ",", header =T)
head(auto)
```
```{r}
potato <- read.table('~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/potatoes.txt', sep="\t", header=T)
head(potato)
```
#### read.csv2() for reading csv files
`Note:` read.csv2() is used to read only comma separated data and does not use the parameter sep

```{r}
other_csv <- read.csv2("~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/who_suicide_statistics_semi.csv", header = T)

head(other_csv)
```
#### using read.delim() for tsv files
`NB:` read.delim() works for tab separated files
```{r}
delim <- read.delim('~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/potatoes.txt', header=F)
head(delim)
```
#### package readr
a package called readr() is 10x faster but it does not come default with R. It has been installed unlike the base functions
Base functions for reading file include:
- read.delim()
- read.table()
-read.csV2()

#### Importing excel files 
> read_excel() is not a base R function. It has to be imported into r as a package.
> `NB:` The col_names here also refers to header as in other file reader functions
> sheet paramter refers to the shhet we want to load. We cannot load more than one sheet at a time

```{r}
library(readxl)
```
```{r}
sht1 <-read_excel('~/Documents/Semicolon/R/10_weeks_of_data_science_with_R/datasets/data/Employees.xlsx', col_names = T, sheet = 1)
head(sht1)
```
#### Importing data from Databases
```{r}
library(DBI)
library(RMySQL)
```

```{r}
host =  'courses.csrrinzqubik.us-east-1.rds.amazonaws.com' 
connect <- dbConnect(RMySQL::MySQL(), dbname = "tweater",
host = host, port = 3306, user = "student", password ="datacamp")
```
> listing tables from databases

```{r}
tables <- dbListTables(connect)
tables
```
> we can use the dbReadTable() function to import data from the database tables.

```{r}
user<-dbReadTable(connect, "users")
user
```

> writing databse object to files

```{r}
write.csv(user, file='user.csv')
```

#### Importing data from statistical packages


#### Baisc Visualizations in R

> Scatter PLots
>> Calculating correlation `NB:` house data was chosen

```{r}
dt<-read.csv(file.choose(), header = T)
corr<-cor(dt$SQFT, dt$Price)
corr
```
> To check the general information/structure of your data, you use str()

```{r}
str(dt)
```



##### Histogram

> Used to plot the relationship between two different numeric data

```{r}
hist(dt$Price, main = "Histogram of Price", xlab = "Price of House")
```
> Using log to normalise a distribution. The eseence of the log normalisation is for modelling strictly . 
>>`NB`: Interpretationn should never be based on the result of log normalisation

```{r}
hist(log(dt$Price), main = "Histogram of Price", xlab = "Price of House")
```
#### Boxplot
> Used to plot the relationship between a numeric varibale and a categorical variable
>> Ensure you are plotting boxplots with no more than 5 categorical variables
>> We have to convert LOcation  or Status features to Factor form Character

```{r}
dt$Location <- as.factor(dt$Location)
dt$Status <- as.factor(dt$Status)
```



```{r}
house_box<-boxplot(dt$Price~dt$Status, xlab = "Status", ylab = "Price", main="House Prices", las=T)
house_box<-boxplot(Price~Status,data = dt, xlab = "Status", ylab = "Price",  main="House Prices")
 
# OR

#house_box
```

```{r}

```

